---
title: "STAT 380 Taryn"
output: html_notebook
---

# Libraries 
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
```

# Dataset
```{r}
SpotifyDataset <- read.csv(file = "SpotifyDataset.csv", stringsAsFactors = FALSE)

colnames(SpotifyDataset)

# Edit this to the columns we want to use
SpotifyDataset <- SpotifyDataset%>%
  select(Track.Name, Artist.Name.s., Popularity, Artist.Genres, Danceability, Energy, Key, Loudness, Speechiness, Acousticness, Instrumentalness, Liveness, Valence, Tempo)

colnames(SpotifyDataset)

SDNums <- SpotifyDataset%>%
  select(Popularity,Danceability, Energy, Key, Loudness, Speechiness, Acousticness, Instrumentalness, Liveness, Valence, Tempo)
```

# By Genre
```{r}
split <- strsplit(as.character(SpotifyDataset$Artist.Genres), ",")

first <- sapply(split, function(x) trimws(x[1]))

SpotifyDataset$Genre <- first

sd1 <-sdMorePop <- SpotifyDataset%>%
  group_by(Genre)%>%
  summarise(count = n()) 

sdMorePop <- SpotifyDataset%>%
  group_by(Genre)%>%
  summarise(count = n())%>%
  filter(count >= 20)
# bc 10000/502

sdMorePop2 <- right_join(SpotifyDataset, sdMorePop)
sdMorePop2 <- sdMorePop2%>%
  select(Track.Name, Artist.Name.s., Genre, Popularity,Danceability, Energy, Key, Loudness, Speechiness, Acousticness, Instrumentalness, Liveness, Valence, Tempo)

sdLessPop <- SpotifyDataset%>%
  group_by(Genre)%>%
  summarise(count = n())%>%
  filter(count < 20)
# bc 10000/502

sdLessPop2 <- right_join(SpotifyDataset, sdLessPop)
```

##  Test
5 + songs per genre and 50 + Median POp
```{r}
test <- SpotifyDataset%>%
  group_by(Genre)%>%
  summarise(count = n(), medianPop = median(Popularity))%>%
  filter(medianPop > 50, count > 5)

test1<- right_join(SpotifyDataset, test)

test2 <- test1 %>%
  select(Track.Name, Artist.Name.s., Genre, Popularity,Danceability, Energy, Key, Loudness, Speechiness, Acousticness, Instrumentalness, Liveness, Valence, Tempo)

test2
```

### summary stats
```{r}
stats <- function(data, names) {
  for (variable in names) {
    variableData <- data[, variable, drop = FALSE]
    print(
      summary(variableData)
    )
  }
}

variables <- colnames(SDNums)
stats(test2, variables)
```

### his
```{r}
histograms <- function(data, names) {
  for (variable in names) {
    variableData <- data[, variable, drop = FALSE]
    
    print(ggplot(
      data = variableData,
      mapping = aes(x = !!as.name(variable))) +
        geom_histogram(color = "black", fill = "salmon") +
        labs(title = paste("Boxplot of", variable),
             x = variable, y = "Frequency") 
    )
  }
}
histograms(test2, variables)
```


### pairs/cors
```{r}
nums1 <- test2%>%
  select(Popularity,Danceability, Energy, Key, Loudness)

nums2 <-test2%>%
  select(Popularity, Speechiness, Acousticness, Instrumentalness, )

nums3 <- test2%>%
  select(Popularity, Liveness, Valence, Tempo)

#pairs(SDNums)
pairs(nums1)
pairs(nums2)
pairs(nums3)

cor(nums1, use = "complete.obs")
cor(nums2, use = "complete.obs")
cor(nums3, use = "complete.obs")
```


## Genre EDA for more popular songs
### boxplots
```{r}
boxplots_by_genre <- function(data, genres, pop) {
  unique <- unique(data[[genres]])
  
  for (genre in unique) {
    genre_data <- data[data[[genres]] == genre, ]
    
    print(ggplot(
      data = genre_data,
      mapping = aes(x = 1, y = !!as.name(pop))
    ) +
      geom_boxplot() +
      labs(title = paste("Boxplot of", pop, "for", genre),
           x = genre, y = pop) +
      theme(axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank())
    )
  }
}

boxplots_by_genre(sdMorePop2, "Genre", "Popularity")
```

### histrograms
```{r}
histograms <- function(data, names) {
  for (variable in names) {
    variableData <- data[, variable, drop = FALSE]
    
    print(ggplot(
      data = variableData,
      mapping = aes(x = !!as.name(variable))) +
        geom_histogram(color = "black", fill = "salmon") +
        labs(title = paste("Boxplot of", variable),
             x = variable, y = "Frequency") 
    )
  }
}


variables <- colnames(SDNums)
histograms(sdMorePop2, variables)
```
### pairs/cor
```{r}
nums1 <- sdMorePop2%>%
  select(Popularity,Danceability, Energy, Key, Loudness)

nums2 <- sdMorePop2%>%
  select(Popularity, Speechiness, Acousticness, Instrumentalness, )

nums3 <- sdMorePop2%>%
  select(Popularity, Liveness, Valence, Tempo)

#pairs(SDNums)
pairs(nums1)
pairs(nums2)
pairs(nums3)

```


### summary stats
```{r}
stats(sdMorePop2, variables)
```


## Genre EDA for more less pop songs
### boxplots
```{r}
boxplots_by_genre <- function(data, genres, pop) {
  unique <- unique(data[[genres]])
  
  for (genre in unique) {
    genre_data <- data[data[[genres]] == genre, ]
    
    print(ggplot(
      data = genre_data,
      mapping = aes(x = 1, y = !!as.name(pop))
    ) +
      geom_boxplot() +
      labs(title = paste("Boxplot of", pop, "for", genre),
           x = genre, y = pop) +
      theme(axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank())
    )
  }
}

boxplots_by_genre(sdLessPop2, "Genre", "Popularity")
```

### summary stats
```{r}
stats(sdLessPop2, variables)
```

# General Data
## Summary Statistics 
```{r}
stats(SDNums, variables)
```

## Boxplots 
```{r}
boxplots <- function(data, names) {
  for (variable in names) {
    variableData <- data[, variable, drop = FALSE]
    
    # Create a boxplot using ggplot2
    print(ggplot(
      data = variableData,
      mapping = aes(y = !!as.name(variable))) +
        geom_boxplot() +
        labs(title = paste("Boxplot of", variable),
             x = "", y = variable) 
    )
  }
}

boxplots(SDNums, variables)
```

## Histograms
```{r}
histograms <- function(data, names) {
  for (variable in names) {
    variableData <- data[, variable, drop = FALSE]
    
    print(ggplot(
      data = variableData,
      mapping = aes(x = !!as.name(variable))) +
        geom_histogram(color = "black", fill = "salmon") +
        labs(title = paste("Histogram of", variable),
             x = variable, y = "Frequency") 
    )
  }
}

histograms(SDNums, variables)
```
There are many songs with a popularity of 0, why is that? Key is not symmetrical. 



## Observations for individual factors
*Popularity:* There are a lot of songs with a popularity of 0, not sure why that is, but it makes it skewed. Without that, it would be fairly symmetrical.   
*Danceability:* Pretty symmetrical with a center around .60-.61    
*Energy:* Fairly symmetrical, slightly skewed left with center around .68-.71.
*Key:* Numbers wise, the mean and median are very similar 5.168 and 5, respectively. When looking at the graph, it does not have a distinct shape, with the a key of 3 being the least chosen and 0 being the most.   
```{r}
SDNums%>%
  group_by(Key)%>%
  summarise(count = n())%>%
  arrange(desc(count))
```

*Loudness:* Pretty symmetrical with center around -.72 - -.65.    

*Speechiness:* Heavily skewed right. With mean of 0.06 and median of 0.04.   
*Acousticness:* Heavily skewed right. With mean of .21 and median of .096.   
*Instrumentalness:* Heavily skewed right. With mean of 0.029 and median of 0.0000057.   
*Liveness:*Heavily skewed right. With mean of .18 and median of .12. 
*Valence:* Numbers wise it is symmetrical but it is more box/square shaped with center around .58.   
*Tempo:* Fairly symmetrical with center around 120-121.  


## Scatterplot
Factors on the y axis, popularity on the x axis. 
```{r}
scatter <- function(data, names) {
  pop <- colnames(data)[sapply(data, is.numeric)][1]  
  for (variable in names) {
    variableData <- data[, c(variable, pop), drop = FALSE]
    
    print(ggplot(
      data = variableData,
      mapping = aes(x = !!as.name(variable), y = !!as.name(pop))) +
        geom_point() +
        labs(title = paste("Scatter Plot of", variable),
             x = variable, y = pop) 
    )
  }
}

scatter(SDNums, variables)

```
No clear patterns or correlation between popularity and a variable. 

## Pairs/Cor
Had to split it up into 3 because it was too small. All have population though
```{r}
nums1 <- SpotifyDataset%>%
  select(Popularity,Danceability, Energy, Key, Loudness)

nums2 <- SpotifyDataset%>%
  select(Popularity, Speechiness, Acousticness, Instrumentalness, )

nums3 <- SpotifyDataset%>%
  select(Popularity, Liveness, Valence, Tempo)

#pairs(SDNums)
pairs(nums1)
pairs(nums2)
pairs(nums3)

cor(SDNums, use = "complete.obs")
```
Popularity does not really have a pattern with any variable. The only variables with a slight linear pattern are energy and loudness. Could be more but they are split up. 
